Evaluation of +S2 sea ice forecast and some other better title

### 2025-04-18 analysis/paper/access-ice.tex
Transfers all comments

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage[inline]{trackchanges}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Evaluation of +S2 sea ice forecast and some other better title},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={red},
  pdfcreator={LaTeX via pandoc}}

@@

\firstpage{1}
\maketitle
\begin{abstract}
The abstract goes here. It can also be on \emph{multiple lines}.
\end{abstract}
\copyrightstatement{The author's copyright for this publication is
transferred to institution/company.}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, interior hidden, sharp corners, boxrule=0pt, enhanced, breakable]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Unlike Arctic sea ice, which has been steadily retreating at least since
the start of satellite records in the early 80s, Antarctic sea ice has
experienced a slightly increasing trend that models systematically
failed to reproduce and puzzled researchers \citep{hobbs2016}. Then, in
2016 Antarctic sea ice extent dropped precipitously and has been at low
and record low levels since, highlighting even more our lack of
understanding of Antarctic sea ice variability and change
\citep{espinosa2024}.

Antarctic sea ice is in a remote location but the potential impacts of
its variability extends far into the lower latitudes, affecting the
circulation of both the ocean and the atmosphere, ocean carbon uptake
and biological processes. It's likely that a reduction in sea ice extent
would reduce the meridional temperature gradient which in turn would
reduce the strength and location of the eddy-driven jet (and vice versa
for an increase in sea ice), which is a crucial component of weather
events. In the northern hemisphere, it has been hypothesised that the
reduction in Arctic sea ice has lead to a weaker, wavier jet, increasing
the frequency of extreme events \citep{barnes2015}. There is some
evidence of this effect in the Southern Hemisphere. Models show a
weakened vortex in the middle atmosphere, increased vertical wave flux
and greater zonal wave 1 amplitude in Springs following a low sea ice
maximum. But models don't fully agree on the exact phase of the zonal
wave response and observations don't show a clear signal
\citep{rea2024}. The jet response to sea ice loss is also very model
dependent, particularly sensitive to its the climatological location and
relatively small compared with the direct effect of CO2 increase and Sea
Surface Temperature (SST) warming \citep{ayres2019}. On climate scales,
it is thought that the temperature signal of a reduction in Antarctic
sea ice mirrors a ``mini global warming'' and experiments also predict a
weakening and northerly shift of the eddy-driven
\citep{raphael2011, england2018, england2020, ayres2022}.

Correctly modelling Antarctic sea is not only necessary for process
understanding and climate projections to inform adaptation strategies.
Accurate seasonal to subseasonal forecast are crucial for operations in
and around the Antarctic continent such as scientific missions,
fisheries and tourism \citep[ \citet{wagner2020}]{desilva2020}. However,
producing accurate Antarctic sea ice forecasts has been challenging due
to model biases and inherent large variability and complexity and it has
lagged behind Arctic sea ice forecasts \citep{zampieri2019}. Dynamical
seasonal forecasts of summer Antarctic sea ice have been shown to
perform worse than relatively simple statistical methods
\citep{massonnet2023}, which also underscores the need for better
understanding of sea ice dynamics.

In this work we evaluate sea ice forecasts produced by ACCESS-S2, the
Australian Bureau of Meteorology subseasonal-to-seasonal prediction
system. This evaluation will inform future work with the model as a
reseach tool and explore the potential of using it as a sea ice
forecasting tool.

\hypertarget{data-and-methods}{%
\section{Data and methods}\label{data-and-methods}}

\hypertarget{s2}{%
\subsection{ACCESS-S2}\label{s2}}

ACCESS-S2 \citep{wedd2022} is the Australian Bureau of Meteorology
current subseasonal-to-seasonal prediction system which replaced
ACCESS-S1 \citep{hudson2017} in October 2021. Both ACCESS-S2 and
ACCESS-S1 consist on the Global Atmosphere 6.0 (GA6)
\citep{williams2015, waters2017}, Global Land 6.0
\citep{best2011, waters2017}, Global Ocean 5.0
\citep{madec2013, megann2014} and Global Sea Ice 6.0 {[}CICE;
\citet{rae2015}{]} model components. The atmosphere has a N216
horizontal resolution (\textasciitilde60km in the mid-latitudes) with 85
levels. The land model uses the same horizontal grid with 4 soil levels.
The ocean component has a 1/4º resolution with 75 vertical levels. The
ice component --based on CICE version 4.1-- has the same resolution than
the ocean and 5 sea ice categories as well as an open water category.

Both systems take atmospheric initial conditions derived from
ERA-interim \citep{dee2011} for the hindcast and from the Bureau's
operational analysis for real-time forecasts. The main difference
between the two are the ocean and ice initial conditions. ACCESS-S1
initial conditions come from the Met Office FOAM system, which uses a
multivariate, incremental three-dimensional variational (3D-Var),
first-guess-at-appropriate-time (FGAT) data assimilation scheme
\citep{waters2015} and assimilates sea surface temperature (SST), sea
surface height (SSH), in situ temperature and salinity profiles, and sea
ice concentration. ACCESS-S2, instead, runs from initial conditions
generated by the in-house data assimilation scheme described in
\citet{wedd2022}. This scheme is a weakly coupled ensemble optimal
interpolation method and assimilates temperature and salinity profiles
from EN4 \citep{good2013} for the hindcast and from the WMO Global
Telecommunication System (GTS) and both the Coriolis and USGODAE Global
Data Assembly Centers (GDACs) for the real-time forecast. SSTs are
nudged to Reynolds OISSTv2.1 \citep{reynolds2007} for the hindcast and
to the Global Australian Multi-Sensor SST Analysis (GAMSSA; Zhong and
Beggs 2008) for the real-time forecast in areas where SSTs are over 0ºC.
Relevant for this work, sea ice concentrations are not assimilated.

@@

\hypertarget{verification-datasets}{%
\subsection{Verification datasets}\label{verification-datasets}}

For verification we use satellite-derived estimates of sea ice
concentration, which estimates the proportion of each grid area that is
covered with ice. Different datasets derived using different algorithms
provide different estimates, each with its own biases and uncertainties.
\citet{meier2019} estimated that the inter-product uncertainty of sea
ice extent is of the order of 1 million km\textsuperscript{2}. As will
be shown below, this spread is minimal compared with the errors in the
ACCESS-S2 and ACCESS-S1 forecasts, so the conclusions are independent of
the dataset used.

We use NOAA/NSIDC's Climate Data Record V4 {[}CDR; \citet{meier2014}{]}
as the primary verification dataset. It takes the maximum value of the
NASA Team \citep{cavalieri1984} and NASA Bootstrap \citep{comiso2023}
sea ice concentration products to reduce their low concentration bias
\citep{meier2014, meier2021}. Both source algorithms use data from the
Scanning Multichannel Microwave Radiometer (SMMR) on the Nimbus-7
satellite and from the Special Sensor Microwave/Imager (SSM/I) sensors
on the Defense Meteorological Satellite Program's (DMSP) -F8, -F11, and
-F13 satellites. The data has a resolution of 25 by 25km and daily from
1978 onwards.

The European Organisation for the Exploitation of Meteorological
Satellites (EUMETSAT) Ocean and Sea Ice Satellite Application Facility
(OSI SAF) \citep{OSISAF} is another satellite-derived sea ice
concentration product. It is based on mostly the same sensors as the NOA
CDR but computed independently using different algorithms. Figures
prepared with this dataset are provided in the supplementary material.

\hypertarget{error-measures}{%
\subsection{Error measures}\label{error-measures}}

@@

Sea Ice Extent is defined as the area of ocean covered with at least
15\% ice. This threshold is motivated by the limitations in satellite
retrieval, which is increasingly unreliable for lower sea ice
conditions.

Pan-Antarctic Sea Ice Extent serves as a rough global measure of the
amount of sea ice but it doesn't take into account the spatial
distribution. A model could have relatively accurate extent of ice but
with different distributions. To account for location errors, we
computed the Root Mean Squared Error (RMSE) of sea ice concentration
anomalies and the Integrated Ice Edge Error {[}IIEE;
\citet{goessling2016}{]}.

We compute RMSE as the square root of the average squared differences
between forecasted and observed sea ice concentration anomalies. We
compute a pan-Antarctic RMSE by averaging over the whole domain, and
also a zonally-varying RMSE by averaging over 24 15° longitude slices
around Antarctica.

The IIEE is defined as the area in which the model misspredicts sea ice
concentration being above or belong 15\% ice. That is, dichotomise sea
ice concentration into areas with more and less than 15\% sea ice both
in the forecast and observations; the IIEE is the area in which forecast
and observations differ.

@@

\hypertarget{results-and-discussion}{%
\section{Results and discussion}\label{results-and-discussion}}

\hypertarget{reanalysis}{%
\subsection{Reanalysis}\label{reanalysis}}

\hypertarget{bias}{%
\subsubsection{Bias}\label{bias}}

@@

{\centering \includegraphics{access-ice_files/figure-pdf/fig-hindcast-extent-1.pdf}

@@

\end{figure}

Figure~\ref{fig-hindcast-extent} a shows the median sea ice extent (left
column) and median difference with respect of NOAA/NSIDC CDRV4 for the
ACCESS-S2 and ACCESS-S1 hindcasts. Median extent of initial conditions
at the first of every month are indicated with circles, while median
extent at the same date but forecasted with the largest lead time
possible for each model (between 274 and 277 days for ACCESS-S2 and
between 213 and 216 days for ACCESS-S1). At this large lead time, the
the information of the initial conditions is essentially lost and the
forecast reverts to each model's preferred equilibrium state.

Both models' equilibrium state (triangles) show low bias of sea ice
extent, particularly in the late-autumn and winter months. This is due
primarily to faster melt during a longer melt season between January and
March and slower growth during March and April. This is then partially
balanced with faster growth between May and July
(Fig.~\ref{fig-mean-growth}). Many sea ice models exhibit this
systematic underestimation during the sea ice minimum and early freezing
season \citep{massonnet2023} and could indicate problems in the
representation of thermodynamics in the model \citep{zampieri2019}. It
is not surprising that both forecasting systems converge to a similar
climatology because they share the same model configuration.

ACCESS-S2 initial conditions (circles) also have a low extent bias,
especially in the late summer-early autumn, while ACCESS-S1 initial
conditions are very close to observations. The faster growth between May
and July reduces the low bias in ACCESS-S2 for forecasts initialised
from April to June, but it adds a high bias in ACCESS-S1.

The difference between the initial conditions and the model preferred
state can be attributed to the effect of data assimilation, which in
ACCESS-S2 is due solely to atmospheric and oceanic data assimilation.
From June to October, in ACCESS-S2 circles are closer to observations
than to the triangles, indicating that the information from the ocean
and atmosphere data assimilation is affecting sea ice and improving the
initial conditions. The rest of the year there is little if any
difference between circles and triangles in ACCESS-S2, indicating that
almost no data assimilation is taking place and the sea ice component of
the model is virtually free-running.

@@

{\centering \includegraphics{access-ice_files/figure-pdf/fig-mean-growth-1.pdf}

@@

\caption{\label{fig-mean-growth}Median daily sea ice extent growth of
ACCESS-S1 and ACCESS-S2 hindcasts and observations. Values are smoothed
with a 2-degree loess smooth with a 30 day window.}

\end{figure}

@@

{\centering \includegraphics{access-ice_files/figure-pdf/fig-bias-1.pdf}

@@

\end{figure}

Figure~\ref{fig-bias} shows the difference of monthly mean sea ice
concentrations between NOAA/NSIDC CDRV4 and ACCESS-S2 hindcasts at the
shortest lag. From October to May, the model underestimates sea ice
concentrations pretty much everywhere there is ice except for the deep
Weddell Sea in April and May, where sea ice concentrations saturate to 1
both in the observations and forecasts. In winter, the differences are
mostly on the sea ice edge, with light positive bias in the African
sector of East Antarctica and negative bias around the Indian Ocean
sector which partially compensate, resulting in the near-zero extent
bias seen in those months (Fig.~\ref{fig-hindcast-extent}).

\hypertarget{anomalies}{%
\subsubsection{Anomalies}\label{anomalies}}

@@

{\centering \includegraphics{access-ice_files/figure-pdf/fig-extent-anom-1.pdf}

@@

\caption{\label{fig-extent-anom}Sea ice extent anomalies for ACCESS-S1
and ACCESS-S2 (black) and NOAA/NSIDC CDRV4 (blue).}

\end{figure}

Figure~\ref{fig-extent-anom} shows monthly sea ice extent anomalies
forecasted at selected lags. The anomalies in this case were computed
with respect of the climatology of each lag, which works as a
bias-correction method. Compared with ACCESS-S1, ACCESS-S2 anomaly
forecast is relatively poor even in the first month, which stays
relatively skilful even at lag 3. ACCESS-S2 shows much bigger
variability than observations, with dramatic lows between 1995 and 2007
and highs between 2007 and 2015.

>>>>

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

\usepackage[inline]{trackchanges}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Evaluation of +S2 sea ice forecast and some other better title},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={red},
  pdfcreator={LaTeX via pandoc}}

@@

\firstpage{1}
\maketitle
\begin{abstract}
The abstract goes here. It can also be on \emph{multiple lines}.
\end{abstract}
\copyrightstatement{The author's copyright for this publication is
transferred to institution/company.}

\section{Introduction}\label{introduction}

Unlike Arctic sea ice, which has been steadily retreating at least since
the start of satellite records in the early 80s, Antarctic sea ice has
experienced a slightly increasing trend up to 2015, which models
systematically failed to reproduce and puzzled researchers
\citep{hobbs2016}. Then, in 2016 Antarctic sea ice extent dropped
precipitously \citep{turner2017a} and has been at low and record low
levels since \citep{purich2023, hobs2024, gilbert2024, josey2024}, even
more highlighting our lack of understanding of Antarctic sea ice
variability and long-term change \citep{espinosa2024}.

Antarctic sea ice is in a remote location, but the potential impacts of
its variability extends far into the lower latitudes, affecting the
circulation of both the ocean and the atmosphere, ocean carbon uptake
and biological processes. It is likely that a reduction in sea ice
extent would reduce the meridional temperature gradient which, in turn,
would reduce the strength and location of the eddy-driven jet (and the
opposite circulation changes with an increase in sea ice), which is a
crucial control on extratropical weather. In the northern hemisphere
(NH), it has been hypothesised that the reduction in Arctic sea ice has
led to a weaker, wavier jet, increasing the frequency of extreme events
\citep{barnes2015}. There is some evidence of this effect in the
southern hemisphere (SH). Models show a weakened vortex in the middle
atmosphere, increased vertical wave flux and greater zonal wave 1
amplitude in austral spring following a low sea ice maximum. But models
do not fully agree on the exact phase of the zonal wave response, and
observations do not show a clear signal \citep{rea2024}. The jet
response to sea ice loss is also very model dependent, particularly
sensitive to its climatological location, and relatively small compared
with the direct effect of CO2 increase and sea surface temperature (SST)
warming \citep{ayres2019}. On climate scales, it is thought that the
temperature signal of a reduction in Antarctic sea ice mirrors a ``mini
global warming'', and experiments also predict a weakening and
equatorward shift of the SH eddy-driven jet
\citep{raphael2011, england2018, england2020, ayres2022}.

Correctly modelling Antarctic sea is not only necessary for process
understanding and climate projections to inform adaptation strategies
but also for accurate seasonal to sub-seasonal forecasts, which are
crucial for operations in and around the Antarctic continent such as
scientific missions, fisheries and tourism \citep[
\citet{wagner2020}]{desilva2020}. However, producing accurate Antarctic
sea ice forecasts has been challenging due to model biases, inherent
large variability and complexity, and it has lagged behind Arctic sea
ice forecasts \citep{zampieri2019}. Dynamical seasonal forecasts of
summer Antarctic sea ice have been shown to perform worse than
relatively simple statistical methods \citep{massonnet2023}, which also
underscores the need for better understanding of sea ice dynamics.

In this work we evaluate sea ice forecasts produced by the Australian
Community Climate and Earth System Simulator -- seasonal version2
(ACCESS-S2), which is the Australian Bureau of Meteorology (BoM)'s
current operational subseasonal-to-seasonal prediction system, and
compare its forecast skill with that of the former system ACCESS-S1.
This evaluation will inform future work with the prediction system as a
research tool and explore the potential of using its sea-ice forecasts
for decision-making.

\section{Data and methods}\label{data-and-methods}

\subsection{ACCESS-S2}\label{s2}

ACCESS-S2 \citep{wedd2022} became operational in October 2021 by
replacing the ACCESS-S1 system \citep{hudson2017}.. The model components
of both ACCESS-S2 and ACCESS-S1 consist of the Global Atmosphere 6.0
(GA6) \citep{williams2015, waters2017}, Global Land 6.0
\citep{best2011, waters2017}, Global Ocean 5.0
\citep{madec2013, megann2014} and Global Sea Ice 6.0 {[}CICE;
\citet{rae2015}{]}. The atmosphere has a N216 horizontal resolution
(\textasciitilde60km in the mid-latitudes) with 85 vertical levels. The
land model uses the same horizontal grid with 4 soil levels. The ocean
component has a 1/4º resolution with 75 vertical levels. The sea ice
component, based on CICE version 4.1, has the same resolution as the
ocean and 5 sea ice categories as well as an open water category.

Both systems take atmospheric initial conditions derived from
ERA-interim \citep{dee2011} for their hindcasts and from the Bureau's
operational analysis for real-time forecasts. The main difference
between the two are the ocean and sea ice initial conditions. ACCESS-S1
ocean initial conditions come from the Met Office FOAM system, which
uses a multivariate, incremental three-dimensional variational (3D-Var),
first-guess-at-appropriate-time (FGAT) data assimilation scheme
\citep{waters2015} and assimilates sea surface temperature (SST), sea
surface height (SSH), in situ temperature and salinity profiles, and sea
ice concentration. ACCESS-S2, on the other hand, runs from initial
conditions generated by the BoM data assimilation scheme described in
\citet{wedd2022}. This scheme is a weakly coupled ensemble optimal
interpolation method and assimilates temperature and salinity profiles
from EN4 \citep{good2013} for the hindcasts and from the WMO Global
Telecommunication System (GTS) and the Coriolis and USGODAE Global Data
Assembly Centers (GDACs) for the real-time forecasts. SSTs are nudged to
Reynolds OISSTv2.1 \citep{reynolds2007} for the hindcasts and to the
Global Australian Multi-Sensor SST Analysis (GAMSSA; Zhong and Beggs
2008) for the real-time forecasts in areas where SSTs are over 0ºC.
Relevant for this work, sea ice concentrations are not assimilated in
ACCESS-S2.

@@

\subsection{Verification datasets}\label{verification-datasets}

For verification we use satellite-derived estimates of sea ice
concentration, which estimates the proportion of each grid area that is
covered with ice. Different datasets derived using different algorithms
provide different estimates, each with their own biases and
uncertainties. \citet{meier2019} estimated that the inter-product
uncertainty of sea ice extent is of the order of 1 million \(km^2\). As
will be shown below, this spread is minimal compared with the typical
errors in the ACCESS-S2 and ACCESS-S1 forecasts, so the conclusions are
independent of the dataset used.

We use NOAA/NSIDC's Climate Data Record V4 {[}CDR; \citet{meier2014}{]}
as the primary verification dataset. It takes the maximum value of the
NASA Team \citep{cavalieri1984} and NASA Bootstrap \citep{comiso2023}
sea ice concentration products to reduce their low concentration bias
\citep{meier2014, meier2021}. Both source algorithms use data from the
Scanning Multichannel Microwave Radiometer (SMMR) on the Nimbus-7
satellite and from the Special Sensor Microwave/Imager (SSM/I) sensors
on the Defense Meteorological Satellite Program's (DMSP) -F8, -F11, and
-F13 satellites. The data has a resolution of 25 by 25 km and daily from
1978 onwards.

The European Organisation for the Exploitation of Meteorological
Satellites (EUMETSAT) Ocean and Sea Ice Satellite Application Facility
(OSI SAF) \citep{OSISAF} is another satellite-derived sea ice
concentration product. It is based on mostly the same sensors as the
NOAA CDR but computed independently using different algorithms. Figures
prepared with this dataset are provided in the supplementary material.

\subsection{Error measures}\label{error-measures}

@@

Sea-ice extent is defined as the area of ocean covered with at least
15\% ice. This threshold is motivated by the limitations in satellite
retrieval, which is increasingly unreliable for lower sea ice conditions
(cite -- probably something from NSIDC).

Pan-Antarctic (net) sea ice extent serves as a rough hemispheric measure
of the amount of sea ice, but it does not take into account the spatial
distribution. A model could have relatively accurate extent of the net
ice but with different regional distributions. To account for location
errors, we computed the Root Mean Squared Error (RMSE) of sea ice
concentration anomalies and the Integrated Ice Edge Error {[}IIEE;
\citet{goessling2016}{]}.

We compute RMSE as the square root of the average squared differences
between forecasted and observed sea ice concentration anomalies. We
compute a pan-Antarctic RMSE by averaging over the whole NOAA/NSIDC
CDRV4 southern hemisphere domain, and also a zonally-varying RMSE by
averaging over twenty-four 15° longitude slices around Antarctica.

The IIEE is defined as the area where the model miss-predicts sea ice
concentration being above or below 15\% ice. That is, dichotomise sea
ice concentration into areas with more and less than 15\% sea ice both
in the forecast and observations.

@@

\section{Results and discussion}\label{results-and-discussion}

\subsection{Bias}\label{bias}

@@

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-hindcast-extent-1.pdf}}

@@

\end{figure}%

Figure~\ref{fig-hindcast-extent} a shows the median sea ice extent (left
column) and median difference with respect of NOAA/NSIDC CDRV4 for the
ACCESS-S2 and ACCESS-S1 hindcasts (right column). Median extent of
initial conditions at the first of every month are indicated with
circles, while median extent at the same date but forecasted with the
largest lead time possible for each model (between 274 and 277 days for
ACCESS-S2 and between 213 and 216 days for ACCESS-S1). At this large
lead time, the the information of the initial conditions is essentially
lost and the forecast reverts to each model's preferred equilibrium
state.

Both models' equilibrium state (triangles) show a negative bias of sea
ice extent, particularly in the late-autumn and winter months. This is
due primarily to faster melt during a longer melt season between January
and March and slower growth during March and April. This is then
partially balanced with faster growth between May and July
(Fig.~\ref{fig-mean-growth}). Many sea ice models exhibit this
systematic underestimation during the sea ice minimum and early freezing
season \citep{massonnet2023} and could indicate problems in the
representation of thermodynamics in the model \citep{zampieri2019}. It
is not surprising that both forecasting systems converge to a similar
climatology because they share the same model formulation.

ACCESS-S2 initial conditions (circles) also have a negative bias,
especially in the late summer-early autumn, while ACCESS-S1 initial
conditions are very close to observations. The faster growth between May
and July reduces the negative bias in ACCESS-S2 for forecasts
initialised from April to June, but it adds a positive bias in
ACCESS-S1.

The difference between the initial conditions and the model preferred
state can be attributed to the effect of data assimilation, which in
ACCESS-S2 is due solely to atmospheric and oceanic data assimilation.
From June to October, in ACCESS-S2 circles are closer to observations
than to the triangles, indicating that the information from the ocean
and atmosphere data assimilation is affecting sea ice and improving the
initial conditions. The rest of the year there is little if any
difference between circles and triangles in ACCESS-S2, indicating that
the ocean and atmosphere data assimilation is not affecting seas ice and
that this component of the model is virtually free-running.

@@

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-mean-growth-1.pdf}}

@@

\caption{\label{fig-mean-growth}Median daily sea ice extent growth of
ACCESS-S1 and ACCESS-S2 hindcasts and observations. Values are smoothed
with a 11-day running mean.}

\end{figure}%

@@

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-bias-1.pdf}}

@@

\end{figure}%

Figure~\ref{fig-bias} shows the difference of monthly mean sea ice
concentrations between NOAA/NSIDC CDRV4 and ACCESS-S2 hindcasts at the
shortest lag. From October to May, the model underestimates sea ice
concentrations in most regions except for the inner Weddell Sea in April
and May, where sea ice concentrations saturate to 1 both in the
observations and forecasts. In winter, the differences are mostly on the
sea ice edge with light positive bias in the African sector of East
Antarctica and negative bias around the Indian Ocean sector which
partially compensate, resulting in the near-zero extent bias seen in
those months (Fig.~\ref{fig-hindcast-extent}).

\subsection{RMSE}\label{rmse}

@@

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-extent-anom-1.pdf}}

@@

\caption{\label{fig-extent-anom}Monthly mean sea ice extent anomalies
for ACCESS-S1 and ACCESS-S2 (black) and NOAA/NSIDC CDRV4 (blue)
forecasted at selected lead times. The RMSE during the overlapping
period (1990--2013) is shown on the top left.}

\end{figure}%

Figure~\ref{fig-extent-anom} shows monthly sea ice extent anomalies
forecasted at selected lead times. Compared with ACCESS-S1, ACCESS-S2
anomaly forecast is relatively poor even in the first month, which stays
relatively skilful even at lag 3. ACCESS-S2 shows much bigger
variability than observations, with dramatic lows between 1995 and 2007
and highs between 2007 and 2015.
